---
title: "PISAData Analysis"
author: "Robin"
date: "February 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#PISA Data Exploration by Robin Garrow
#This report explores a dataset containing results from the Program for International Student Assessment(PISA).  The test hopes to assess #education systems worldwide. 
#The analysis below examines how the three particular groups of individuals affect student results. 

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

library(ggplot2)
library(dplyr)
library(lubridate) 
library(gridExtra)
library(tidyr)
library(doBy)
library(corrplot)


```
```{r echo=FALSE, warning=FALSE, Load_the_Data}
# Set the working directory and load the Prosper data
setwd("T:/UdacityDesktop/EffectiveVis/PISAProject/")
pd <- read.csv("pisa2012.csv")
dim(pd)
```


#The original data contains 636 variables and 485,490 records.  

```{r warning=FALSE, subset_data}
# Intentionally display this section in the report so data manipulations can be
# observed.
# subset the PISA Data to capture Canada, United States, and data of interest and exclude cancelled 
#loans

spd <- subset(pd, CNT == "Canada" | CNT == "United States of America",
        select = c(CNT, NC, STIDSTD, ST04Q01, ST35Q02, ST35Q04,
                    ST83Q03, W_FSTUWT, PV1MATH, PV2MATH, PV3MATH, PV4MATH, PV5MATH))

#
#add the mean of the five plausible scores to the dataframe
spd$PVmean <- apply(subset(spd, NC != "", select=c(PV1MATH, PV2MATH, PV3MATH, PV4MATH, PV5MATH)), 1, mean)

#Now use the weight and mean columns to determine the average score for the country and for those who said Agree to the question, "Friends Work #on hard on mathematics". (on report from PSA this is field STP5077 but in the Excel this is ST35Q02)    
friendspd <- spd %>%
  group_by(CNT, ST35Q02) %>%
  summarise(wtdmean = weighted.mean(PVmean, W_FSTUWT, na.rm = TRUE))
 friendspd$grouping <- "Friends"
 friendspd$question <- "My friends work hard in math"
 names(friendspd)[2] <- "Response"

#corresponding question is STP5079 on PISA website
parentspd <- spd %>%
  group_by(CNT, ST35Q04) %>%
  summarise(wtdmean = weighted.mean(PVmean, W_FSTUWT, na.rm = TRUE))
parentspd$grouping <- "Parents"
parentspd$question <- "My parents believe math is important"
 names(parentspd)[2] <- "Response"

#corresponding question for teachers is STP5226
teacherspd <- spd %>%
  group_by(CNT, ST83Q03) %>%
  summarise(wtdmean = weighted.mean(PVmean, W_FSTUWT, na.rm = TRUE))
teacherspd$grouping <- "Teachers"
teacherspd$question <- "My teacher helps with learning"
 names(teacherspd)[2] <- "Response"
#combine the files
outpd <- full_join(friendspd, parentspd)
outpd <- full_join(outpd, teacherspd)


dim(spd)
summary(spd)
#Now keep the reduced file 
write.csv(outpd, "USACAD.csv")
          
```

Analaysis Notes:
See the .png files that come from the PISA site querys.

In Canada, it doesn't appear that students who have friends who work hard
relates to the given students Math scores.  Note how the scores are similar
 523, 521, 520, and 517 for Strongly Agree to Strongly Disagree.  Whereas in the 
 United States a much larger descrepancy exists between Strongly agree vs. 
 strongly disagree(490 vs 453)  This seems to indicate that US students are more
 affected by how their friends do in math than in Canada.  What is it about 
 peer pressure or friend influence in the US that makes such an impact on student
 success?
 
In Canada Students who answered that they "Strongly Agree" 
that their parents believed Math is important scored much better than 
those who answered "Disagree" since US didn't have results for "Strongly Disagree"
(533 vs 488 or a difference of 47 points)
while in the United States the scores were 496 vs 443 for the same response
This is a difference of 35 points which suggests that in both countries 
students who indicate parents believe math is important has large impact
on math scores.

In Canada the difference between Strongly agree vs Strongly disagree is 39 points
(527 vs 488) yet in the US the difference between these two categories is only
11 points(493 vs 482) indicating that when students in the UK perceive their 
teacher helps with learning they do much better than those students do not while 
in the US it seems whether students Strongly agree or Strongly disagree with the same 
statement has nowhere near the impact on student's scores.  What is it saying 
about teacher interest in the US - it's almost that whether or not a student 
believes their teacher is interested in helping with learning has very little
influence on test scores.




```

```
